### 2019-09-20 draft.tex
updated main body text

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subfigure}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{multicol}
\renewcommand{\baselinestretch}{1.5}
\usepackage[utf8]{inputenc}
\usepackage{etoolbox}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{12}{15}\selectfont}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsection}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\usepackage{etoolbox}
\patchcmd{\thebibliography}
  {\settowidth}
  {\setlength{\itemsep}{0pt plus 0.01pt}\settowidth}
  {}{}
\apptocmd{\thebibliography}
  {\small}
  {}{}
\author{Shivesh Pathak}
\title{Developing effective theories for single and bilayer graphene using density matrix downfolding}

\begin{document}
\maketitle
\begin{abstract}
\thispagestyle{empty}
Building accurate model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
A recently developed density matrix downfolding (DMD) \cite{Zheng2017} method provides a direct framework for downfolding the \textit{ab initio} Hamiltonian to a many-body effective Hamiltonian using only \textit{ab initio} calculations. 
Here I propose a thesis plan for developing atomic scale many-body model Hamiltonians for single- and simple bi-layer graphene systems within the DMD framework using high accuracy \textit{ab initio} quantum Monte Carlo (QMC) calculations.
The results of my thesis will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene, and will provide a baseline for a continuing study of the twisted bi-layer graphene (TBLG) system.
To demonstrate my preparedness, I present progress on a many-body model for the CuO molecule and my work in QMC methods development.
\end{abstract}
Committee: Lucas Wagner, Brian DeMarco, Andre Schleife, Karin Dahmen
\clearpage
\pagebreak
\setcounter{page}{1}
\section{Introduction}
Developing model Hamiltonians for materials is a fundamental problem in modern condensed matter physics.
A model Hamiltonian is a simplified description of a material which can accurately reproduce the low-lying eigenstates and energies of that material.
Accurate models give clear and direct indications of effective interactions within a material and underpin modern theoretical understanding of ideas like emergence and symmetry breaking.
On a practical level, calculations related to these low-lying eigenstates can be carried out on the much simpler model and still yield accurate results reflective of reality.

Yet unresolved is a paradigm for building many-body model Hamiltonians which can describe strongly correlated systems.
Band and Fermi liquid theory have been successful in accurately modeling systems where effective interactions are negligible at low energy scales.
These theories, however, fail in describing the low-energy excitations of strongly correlated systems like high Tc superconductors and TBLG.
Minimal many-body models like the Hubbard \cite{Hubbard1963}, Kanamori \cite{10.1143/PTP.30.275}, t-J \cite{Chao_1977} and Heisenberg models have been proposed to describe such systems, their effectiveness in describing realistic materials is unclear.

In an attempt to build accurate many-body Hamiltonians for realistic systems, a framework for model development using solutions from \textit{ab initio} simulations has emerged.
Common approaches involve density functional theory (DFT) \cite{PhysRev.140.A1133, PhysRevLett.87.047003} alongside random phase approximation or other methods \cite{PhysRevB.62.R16219, PhysRevB.70.195104, PhysRevB.88.075106}.
L\"owdin methods and canonical transformations have also been used to develop effective theories \cite{doi:10.1063/1.4802766, PhysRevA.81.013402, doi:10.1063/1.2196410}.
While useful, it is typically impossible to know whether models built from these techniques accurately describe reality, as measures of model quality are rarely reported.
Instead, model validation relies on comparisons to experiment.

The recently developed DMD \cite{Zheng2017} method provides a more direct link between \textit{ab initio} and many-body effective Hamiltonians.
Similar to the methods above, DMD uses information from \textit{ab initio} simulations on the system to fit the best parameters for a given many-body model \textit{ansatz}.
Unlike the methods above, the resultant fit model is accompanied by a clear measure of model quality, quantifying the ability of the model to reproduce the low-energy spectrum of the system.
Since model quality is reported, DMD allows for direct comparison between different model \textit{ansatzes} as well, an ability which can be used to deduce important model terms using \textit{ab initio} calculations.
DMD has been used to develop accurate many-body models for benzene, a 1-d hydrogen chain, and an on-site Hubbard model for graphene \cite{Zheng2017, Wagner2015}.

The utility of DMD is made apparent when studying the recent surge in model development for TBLG.
\textit{Ab initio} model development has so far been restricted to tight-binding models which accurately describe single particle effects like band flattening, but do not address interactions \cite{Fang, Fang2016, PhysRevX.8.031088, PhysRevX.8.031087}.
Minimal interacting models have been proposed using traditional theoretical methods, but without a link to the \textit{ab initio} Hamiltonian it is unclear whether they can describe the real system \cite{PhysRevB.98.045103, PhysRevB.100.125116, PhysRevX.8.031089}.
DMD can bridge the gap between these methods by providing a framework to build an accurate many-body model directly from \textit{ab initio} simulations on TBLG.

Instead of jumping head first into TBLG, however, I propose a two year plan for modeling the AA and AB stacked BLG systems using DMD with \textit{ab initio} fixed-node diffusion Monte Carlo (FN-DMC) calculations.
I will first develop a model for single-layer graphene (SLG) and then extend this to the BLG systems by introducing inter-layer couplings.
The SLG model will provide an accurate evaluation of the effective Coulomb interaction spatial dependence without reliance on analytic forms like Ohno \cite{Ohno1964}, and will describe the role of phonons in modifying these interactions. 
The AA/AB BLG models will address the form and magnitude of inter-layer coupling in the bi-layer, and the effect of a single layer on the intra-layer interactions of its neighbor.
My work will also provide a baseline for follow-up studies on the effects of commensurate twists on inter- and intra-layer coupling in TBLG.

\subsection{Density Matrix Downfolding (DMD)}
The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the Hilbert space $\mathcal{H}$
\begin{equation}
\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + E_\text{ion}(\{R_I\})  
\label{eq:Hab}
\end{equation}
where index $i,j$ correspond to electrons, $I$ to ions, $E_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and energies of $\hat{H}_\text{ab}$.
$\hat{H}_\text{eff}$ is taken to be a linear operator on $\mathcal{LE}$ and can be expressed as a linear combination of Hermitian operators $\hat{d}_k$, with coupling constants $g_k$ and energy offset $E_0$
\begin{equation}
\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
\label{eq:Heff}
\end{equation}
The process of taking $(\hat{H}_\text{ab}, \mathcal{H}) \rightarrow (\hat{H}_\text{eff}, \mathcal{LE})$ is called downfolding of $\hat{H}_\text{ab}$.

DMD provides a framework for this downfolding using energy functionals.
In their seminal paper on DMD, Zheng \textit{et al.} \cite{Zheng2017} show that $\hat{H}_\text{eff}$ will exactly reproduce the low-lying eigenstates of $\hat{H}_\text{ab}$ if
\begin{equation}
\begin{split}
\forall |\Psi\rangle \in \mathcal{LE},\ E_\text{ab}[\Psi] = E_\text{eff}[\Psi].
\end{split}
\label{eq:DMD}
\end{equation}
Here $E_\text{ab}[\Psi] \equiv \langle \Psi | \hat{H}_\text{ab} | \Psi \rangle$ is the known \textit{ab initio} energy functional and $E_\text{eff}[\Psi] \equiv \langle \Psi | \hat{H}_\text{eff} | \Psi \rangle$ is the to-be-determined effective energy functional.
Deviation from equality manifests as an error $\epsilon[\Psi] \equiv E_\text{ab}[\Psi] - E_\text{eff}[\Psi]$ which quantifies the model quality.
In short, the key insight of DMD is that the downfolding $\hat{H}_\text{ab}$ of can be cast to a functional appoximation of $E_\text{ab}[\Psi]$ on $\mathcal{LE}$.

Given the general linear form of $\hat{H}_\text{eff}$ seen in \eqref{eq:Heff}, 
the functional approximation reduces to fitting a linear model that minimizes $\epsilon[\Psi]$ over $\mathcal{LE}$.
The fitting is tackled in three steps.
First, low-energy states are sampled from $\mathcal{LE}$.
Next, a pool of candidate operators $\{\hat{d}_k\}$ which may be included in the final model are selected.
Finally, a linear regression is carried out with the expectation values $\{\langle \Psi | \hat{d}_k|\Psi \rangle, E_\text{ab}[\Psi] \}$  evaluated on the sampled states as the independent and target variables.

Low-energy states are accessed using any wave function based \textit{ab initio} simulation technique, but two broad principles are used for selecting a method.
First, the method should be able to generate accurate low-energy wave functions in the \textit{ab initio} Hilbert space, as the accuracy of the low-energy states dictate the accuracy of the final model.
Second, the method should be relatively inexpensive, as large amounts of training data are preferred for a more robust regression.
QMC methods have been popular choices in previous DMD calculations \cite{Zheng2017, Wagner2015} due to their polynomial scaling computational cost, high accuracy, and applicability to both molecular and lattice systems.

Candidate descriptors are selected using both physical intuition and statistical
methods, seen clearly in a DMD study on the stretched FeSe molecule \cite{Zheng2017}.
An essential first pass involves selecting a large set of descriptors using knowledge of the low-energy degrees of the freedom of the system and physical principles like symmetry.
This set is then further reduced by discarding descriptors which do not exhibit covariance with the energy over the low-energy sampled states.
Note, while there is no restriction on the candidate operators other than Hermiticity, 1- and 2-body second quantized operators are commonly used.

The results of the linear regression are an estimate of the parameter values $g_k$ and model error $\epsilon[\Psi]$.
The error estimate usually takes the form of a goodness of fit, such as a cross-validated $R^2$ score or root mean squared error.
These error estimates present a direct measurement of model quality, ensuring that DMD not only gives the best fit parameters but also an assessment of model validity.
Further, by quantifying model quality through statistical error estimates, it is possible to compare the accuracy of various low-energy \textit{ansatzes}.

\subsection{Fixed-node diffusion Monte Carlo (FN-DMC)}
Diffusion Monte Carlo (DMC) is a QMC method which projects out the ground state of a real-space Hamiltonian given some initial trial wave function.
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}$ with ground state $|\Phi_0\rangle$. 
Applying the projector $e^{-\tau \hat{H}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projects out the ground state as long as the trial wave function is not orthogonal to the ground state. 
DMC provides a stochastic implementation of the projection by moving samples from $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle$, which is approximated using a Trotter expansion $G(R, R^\prime, \tau) \sim \Big[e^{-d\tau(\frac{V(R) + V(R^\prime)}{2} - E_T)} \langle R| e^{-d\tau\hat{T}}|R^\prime \rangle + O(d\tau^2) \Big]^N $ 
where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.

DMC, however, suffers from a fermion sign problem which is alleviated via a fixed-node approximation.
Under the fixed-node approximation the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.

While seemingly disadvantageous, the variational nature of FN-DMC plays
a key role in generating the low-energy states needed for DMD.
Consider a set of trial wave functions with varying nodal structures.
The final projected $\Psi_\text{DMC}$ for these different trial functions will be the lowest energy states in $\mathcal{H}$ for the given nodal structures.
If the initial trial wave functions are appropriately chosen, the final projected states will be within the target $\mathcal{LE}$.
The training data for DMD are then extracted by evaluating the expectation values of the Hamiltonian and candidate operators using the final distributions of walkers from the FN-DMC calculations \cite{Wagner2017}.

\section{Preliminary Work}
\subsection{Non-orthogonal determinants in FN-DMC trial wave functions}
In order to become acquainted with QMC algorithms I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{doi:10.1063/1.5052906}.
The MSJ+NO trial wave functions take the form
\begin{equation}
\Psi=e^{J(\vec{\alpha})} \sum_I e^{\hat{\Theta}(\vec{\theta_I})} C_I |D_I (\{ \phi\})\rangle
\end{equation}
where the determinants $|D_I\rangle$ are composed of single particle orbitals from a self-consistent field calculation, $e^J$ is a spin-independent Jastrow factor, and $e^{\hat{\Theta}}$ is a spin-dependent orbital rotation matrix.
The wave function has variable parameters $\vec{\alpha}, \vec{\theta}, \vec{C}$ which can be optimized.

I assessed the efficiency and compactness of the new function by calculating the ground state energy of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with orthogonal determinants (MSJ+O). 
The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in FN-DMC calculations. 
The average improvement in the FN-DMC energy with the MSJ+O trial wave functions was $\langle E_{MSJ} - E_{MSJ,O}\rangle =$ 0.14 eV with an additional reduction of 0.032 eV when using the MSJ+NO trial wave function.
While the benefit using the MSJ+NO trial wave function seems small, the energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the energy using an MSJ+O trial 
function with 55 determinants.

Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result since introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region. 
The results indicated that using non-orthogonal determinants may lead to more compact trial wave functions for small molecules.

\subsection{Effective theory for CuO molecule using DMD}
To become acquainted with DMD, I constructed a many-body effective model for the CuO molecule with fixed bond length $r_e$ = 1.725 \r{A}.
In the following three sections I will detail the steps I took towards developing a model for the CuO molecule using DMD, followed by a comparison of model spectra to experiment.

\vspace{-10pt}
\paragraph{Sampling low-energy states}
Low-energy states were generated by using FN-DMC projected multi-Slater-Jastrow (MSJ) trial wave functions.
The trial wave functions took the form
\begin{equation}
|\Psi_T(\vec{p}) \rangle =  e^{J}\sum_{j} p_j|\text{D}_j\rangle,
\label{eq:sampling}
\end{equation}
where $e^J$ is a Jastrow factor, $|\text{D}_j\rangle$ are selected determinants, and $p_j$ are variable coefficients.
By varying $\{p_j\}$ one can move around the low-energy space if the set $\{|\text{D}_j \rangle\}$ is representative of the low-energy excitations in the system.
The Jastrow factor explicitly introduces electronic correlation and reduces the localization error within FN-DMC \cite{Foulkes2001}.

The determinants were constructed from the orbitals of symmetry-targeted unrestricted Kohn Sham (UKS) calculations.
Only determinants composed of Cu 3d, 4s, and O 2p orbitals with at most a single hole in the Cu 3d shell were included in \eqref{eq:sampling} to match the low-energy excitations seen in experiment.
The UKS calculations were done using PySCF \cite{Sun2018}, a B3LYP functional \cite{doi:10.1063/1.464304}, Trail-Needs psuedopotential and VTZ Trail-Needs basis \cite{doi:10.1063/1.4811651}.

A shell sampling scheme was used to select coefficients $p_j$ for the trial wave functions.
I began by fixing the coefficient $p_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
For each choice of $w$, n = 5 states were selected by sampling the unassigned coefficients from independent, standard normal distributions such that $\sum_j p_j^2 = 1$. 
This procedure generated a uniform set of samples from spherical shells at decreasing radii in determinant space centered at $|\text{D}_0\rangle$.
The process was repeated by centering at every determinant in \eqref{eq:sampling}, resulting in a sample set of overlapping shells which fill the low-energy space.

A single three-body Jastrow factor was used for all trial wave functions, and was optimized on the lowest energy determinant using an energy optimization scheme \cite{Toulouse2007}.
This choice greatly reduced the computational cost of sampling low-energy states as optimization was not necessary for every sampled state.
The resulting MSJ trial function remained low in energy as variations of the Jastrow factor among low-energy states is expected to be small.

The sampled MSJ trial wave functions were projected using FN-DMC to generate the final set of low-energy sample states.
FN-DMC calculations were conducted with T-moves \cite{PhysRevB.74.161102} to ensure the variational principle holds while using pseudopotentials, and a timestep of $\tau = 0.01$.
All QMC calculations were done in QWalk \cite{Wagner2009}.

\vspace{-10pt}
\paragraph{Selecting candidate descriptors}
The following set of 1- and 2-body operators were selected as candidates for the model :
\begin{equation}
\begin{split}
\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +
t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}\\
  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} \Bigg)  +
J_{sd}\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i} + U_s \hat{n}_{4s,\uparrow}\hat{n}_{4s,\downarrow} + E_0.
\end{split}
\label{eq:models}
\end{equation}
The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in Figure \ref{fig:cuo_regr}a.

The first chunk of descriptors were selected using physical intuition.
As described above, the low-lying excitations in CuO observed in experiment involve the Cu 3d, 4s and O 2p atomic orbitals.
Given that orbitals also hybridize, the occupation energies of and hopping terms between these orbitals were selected as candidate operators.
Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \cite{Data2009}.

The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled states.
Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3 eV, an underestimation of 4 eV.
The effective on-site Coulomb repulsion $U_s$ ensures the model places all such $4s^{2}$ states high in energy.

\begin{figure*}
\centering
%\hspace*{-0.25in}
\includegraphics[width=1.0\linewidth]{./figs/cuo_regr_2.pdf}
\caption{Figures related to model regression for the CuO molecule: a) Intrinsic atomic orbitals used as a single particle basis for the model Hamiltonian, b) Results of PCA showing the CEVR and CN of a data set containing N$_{comp}$  highest ranked PCs, c) The final regressed model parameters, d) Comparison of four lowest lying eigenstates of CuO molecule between experiment and three computation methods.}
\label{fig:cuo_regr}
\end{figure*}

\vspace{-10pt}
\paragraph{Fitting the model}
A direct attempt at fitting all eleven terms in \eqref{eq:models} leads to a high goodness of fit, R$^2$ = 0.99, but unreasonably large coefficients for model.
This is seen in a huge energy gap of $\epsilon_{2p_\pi} - \epsilon_{3d_\delta} = $ 5.87 eV, which is 2 $-$ 3 eV in experiment and the sampled data.
This inflation of model parameters was found to be associated with multicollinearity among the candidate descriptors, quantified by a large condition number (CN) of 1470.

A principal components analysis (PCA) on the descriptors illustrates the multicollinearity in the data set.
In Figure \eqref{fig:cuo_regr}b I show the cumulative explained variance ratio (CEVR) and CN of a matrix composed of the $N_\text{comp}$ highest ranked PCs varying from one to eleven.
The CEVR indicates how much of the variation in the descriptor values a subset of PCs capture, and the CN quantifies the multicollinearity among these PCs.
The sharp decrease in the condition number and sustained high CEVR up to nine PCs indicates the presence of only nine linearly independent variations within the descriptor set.

To reduce multicollinearity while maintaining a high goodness of fit, a model was fit to these nine PCs.
This model yields an $R^2$ = 0.98 with a CN of 257.
An inverse PCA transformation was used to map the fit coefficients corresponding to the PC descriptors back into the original eleven coefficients of \eqref{eq:models}.
These coefficients can be seen in Figure \ref{fig:cuo_regr}c with single standard deviation bootstrap error estimates.
The procedure carried out here is well established and called principal components regression (PCR) \cite{10.2307/2348005}.

\vspace{-10pt}
\paragraph{Solutions of the model}
The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization.
Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
Further, the term symbols assigned in experiment match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.

The energies for the UKS and CCSD(T) \cite{Xian2000} states are also shown in Figure \ref{fig:cuo_regr}d as green and red circles, respectively.
The UKS energies were calculated in this work en route to constructing the determinants $|\text{D}_j\rangle$ used in \eqref{eq:sampling}.
The model energies agree with both theoretical techniques within errorbars for the first three states shown.
The energy of the highest energy $^2\Delta$, however, is severely over estimated using both UKS and CCSD(T), by nearly 1 eV.
This over estimation may arise from an inadequate relaxation of the $\sigma$ type orbitals which form a three orbital hybridization complex in this system.

\vspace{-10pt}
\paragraph{Next steps}
I have shown that my model has comparable accuracy to modern computational techniques.
This poses the question, however: why use DMD if simpler methods are just as accurate?
In the next steps of the project I will address this question by demonstrating the unique interpretability of a model.
I plan on accomplishing this by refitting my model after dropping specific interactions, like $n_{4s,\uparrow} n_{4s,\downarrow}$, and quantitatively evaluating the effect on the model spectra, eigenstates, and renormalization of non-interacting parameters.
In doing so, I will be able to describe the precise effects of certain interactions on the low-energy spectra of CuO, a level of detailed interpretation unavailable to methods like CCSD(T).

\subsection{Methods development and coding}
In addition to directly applying DMD, I have worked on the methods and code development needed to efficiently carry out downfolding.
My first contribution was developing the theory and implementation of an efficient MC estimator for wave function parameter gradients.
The parameter gradient of a wave function, $\langle \Psi(\vec{p}) | \Psi_p (\vec{p}) \rangle$ where $| \Psi_p (\vec{p}) \rangle \equiv \partial_{p} |\Psi (\vec{p}) \rangle$ , is a quantity that pops up frequently in QMC, from optimization algorithms \cite{Toulouse2007} to density matrix downfolding \cite{2019APS..MARF18008W}.
Unfortunately, a naive Monte Carlo estimate of the gradient has an infinite fourth moment due to divergences near the nodal surface of $|\Psi \rangle$, so MC error estimates for the gradient are very unstable.
I proved that the fourth moment can be made finite when using a stable estimator 
\begin{equation} 
\hat{\theta}=\begin{cases}
\frac{\Psi_p(\textbf{r})}{\Psi(\textbf{r})} & \text{away from node} \\
 \frac{\Psi_p(\textbf{r})-\Psi_p(\textbf{N})}{\Psi(\textbf{r})} & \text{near node}.      
\end{cases}
\end{equation}
where $\vec{N}$ is a nodal point.
I implemented, tested and integrated this estimator into QWalk.
A publication with performance details of the estimator is in progress.

I also made significant contributions to a new QMC package called PyQMC. 
This package is oriented towards methods development, and will be the platform for most of my future DMD calculations.
I implemented multi-Slater wave functions, ran consistency checks against QWalk, and worked on execution speedups for costly methods like VMC and wave function optimization.
I plan on implementing 2-body RDM evaluation and Ewald summations too, thus providing the baseline tools for DMD on molecular and periodic systems.

\section{Proposed work}
I propose a two year thesis plan for developing atomic scale many-body
model Hamiltonians for SLG and simple BLG systems within the DMD framework using high accuracy \textit{ab initio} QMC calculations.
To do so, I will be building a sequence of four many-body model Hamiltonians for increasing complex systems: the benzene molecule, SLG, AA stacked BLG, and AB stacked BLG.
The following sections detail the sampling scheme and candidate models that will be used for downfolding each system.

\subsection{Constrained variational Monte Carlo (cVMC)}
A new constrained cVMC method will be used to sample the low-energy wave functions necessary for DMD.
The cVMC method works by minimizing the following cost function for a parameterized wave function $|\Psi(\vec{p})\rangle$ with respect to the parameters $\vec{p}$:
\begin{equation}
\vec{p}^* = \text{argmin} \ \langle \Psi(\vec{p}) | \hat{H}_\text{ab} | \Psi(\vec{p}) \rangle - \sum_k [\lambda_k \langle \Psi(\vec{p}) | d_k - d_{k}^0 | \Psi(\vec{p}) \rangle]^2.
\end{equation}
The quantity $\vec{d}$ is short hand for a list of candidate descriptors to sample variations in, $\vec{d}^*$ is the target values of those descriptors, and $\vec{\lambda}$ controls the importance of finding a state with low-energy versus one with descriptor values near $\vec{d}^*$.
This method allows one to target low-energy wave functions which sit near $\vec{d}^*$ in descriptor space, providing a more systematic method for sampling the low-energy space.

I will use an MSJ parameterization in the cVMC sampling.
The multi-Slater expansion will consist of determinants of single particle orbitals calculated using DFT.
Given the nature of graphene-like systems, as described below, the determinants will include the DFT ground state as well as molecular orbital excitations among the $\pi$ and $\pi^*$ orbitals which form the active space.
A three body Jastrow factor will be applied to ensure that correlations are included in the wave function and that the wave function is low in energy.
The total set of parameters $\vec{p}$ will be Jastrow and determinental coefficients.

\subsection{Benzene}
While relatively simple, the benzene molecule shares many similarities with and forms the basic unit of SLG.
A schematic drawing of the molecule is shown in Figure \ref{fig:proposed}a.
The delocalized $\pi$ orbitals in benzene carry over directly into SLG where they are the primary contributors to the well known Dirac bands.
Interaction effects seen in the benzene molecule are also reflected in SLG, such as on-site double occupation energies for the carbon atoms \cite{Zheng2017, Wagner2015}.

\begin{figure*}
\centering
\includegraphics[width=1.0\linewidth]{./figs/proposed.pdf}
\caption{Schematic diagrams for the a) benzene molecule, b) AB and AA stacked BLG \cite{doi:10.1063/1.5007713}.}
\label{fig:proposed}
\end{figure*}

My objective will be to reproduce a previously developed many-body model for the benzene molecule using QMC DMD \cite{Wagner2015}.
In that study, an extended Hubbard model built on the six $\pi$ orbitals was found to perform well in describing the low-lying excitations of the material.
The final model took the form 
\begin{equation}
H_\text{benz} = -\sum_{\langle i,j \rangle} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{\langle i,j \rangle}V_{ij} n_i n_j + \sum_{\langle \langle i,j \rangle\rangle}V_{ij} n_i n_j + E_0
\label{Hbenz}
\end{equation}
where the brackets denote nearest neighbor and next-nearest neighbor summations.
As such, the model \textit{ansatz} that I will try to fit will be \eqref{Hbenz}.

This first stage of model development provides a playground for benchmarking new model fitting procedures.
Downfolding techniques have advanced significantly since the initial model for benzene was developed in 2015.
For example, methods like constrained VMC for low-energy sampling have been posited, but not thoroughly tested.
By comparing models built using new methods directly to the 2015 model, it will be possible to quantify the efficiency and effectiveness of these methods.

\subsection{Single layer graphene}
SLG is composed of a triangular Bravais lattice with two carbon atoms per unit cell.
In terms of a non-interacting picture, the $\sigma$ orbitals composed of C $2s, 2p_x,$ and $2p_y$ lie below the Fermi level.
The bands which cross the Fermi level, and compose the distinct Dirac cones seen in SLG, are primary $\pi$ bonded molecular orbitals composed of C $2p_z$ orbitals.
Further, reduced screening in SLG leads to long range Coulomb interactions in the system \cite{Elias2012, Yu2013}.

I will work on extending an on-site Hubbard model for SLG \cite{Zheng2017, Wagner2015} to one with long-range Coulomb interactions and electron-phonon coupling.
The candidate model will be of the form:
\begin{equation}
H_\text{SLG} = -\sum_{i,j} t_{ij}c_i^\dagger c_j + U \sum_i n_{i\uparrow}n_{i\downarrow}  + \sum_{i,j} V_{ij} n_i n_j + E_0
\label{Hslg}
\end{equation}
where the indices correspond to the $\pi$ orbitals on the copper atoms.
I will consider both long range hopping and density-density interaction terms extending across the computational unit cell, with the final goal of developing position dependent parameter values $t(r), V(r)$.
Phononic contributions will be incorporated by fitting a sequence of models to altered geometries, leading to displacement dependent parameters similar to a spin-phonon model for the cuprates \cite{Wagner2014}.

An important result from this study would be an accurate evaluation of $V(r)$ directly from \textit{ab initio} simulations.
This result would be able to validate commonly used models for the spatial dependence, like the Ohno form \cite{Ohno1964}.
It will also help develop understanding of interaction effects in graphene which depend on the details of $V(r)$, like edge magentism \cite{PhysRevB.95.195420} and non-local electrodynamics \cite{PhysRevB.90.045137}.
Additionally, by introducing displacement dependent parameters, one can determine whether phonons in graphene serve to simply renormalize interactions, or drastically alter them.
This determination is integral to understanding effects of strain on electronic properties as seen in experiment \cite{Naumis_2017}.

\subsection{Bilayer graphene}
The last systems I will develop models for are AA and AB stacked BLG, shown in Figure \ref{fig:proposed}b.
AA stacked BLG is simply two sheets of graphene placed on top of one another where each atom is aligned with the atom beneath it.
In AB stacked BLG, a carbon atom from the lower layer is present in the center of each carbon hexagon of the upper layer, and vice versa.
As such, one can think of AB BLG as a shifted version of AA BLG, where the shift is half the size of a single carbonic hexagon.

I anticipate that the candidate descriptors should include interlayer couplings as well as the terms seen in SLG.
As such, the candidate model for BLG would be
\begin{equation}
H_\text{BLG} = H_\text{SLG}^{(1)} + H_\text{SLG}^{(2)} - \sum_{i_1, i_2} t_{i_1, i_2}^\prime c_{i_1}^\dagger c_{i_2} + h.c. + E_0
\label{Hblg}
\end{equation}
There are two terms for the single layers from \eqref{Hslg} plus an interlayer hopping given by the term $t_{i_1, i_2}^\prime.$  
The index $i_1$ refers to layer 1 and $i_2$ to layer two.
While this is an initial \textit{ansatz}, additional inter- and intra-layer terms can be included if necessary.

With these two models, one can begin to answer some questions about the inter-layer coupling in BLG.
First, one can determine, at least for the simplest orientations, whether the inter-layer coupling takes the form of a hopping or an interaction term.
Quantitative details of the interlayer coupling spatial dependence, which are intimately related to the strongly correlated behavior in TBLG, can also begin to be addressed.
Further, when combined with the SLG model, it will be possible to determine how the intra-layer model is affected by the presence of a neighboring graphene sheet.
Lastly, the simple BLG models will enable follow-up studies on the twist dependence of model parameters for TBLG.

\vspace{-10pt}
\paragraph{Proposed timeline} I believe that developing effective theories for the benzene molecule, SLG, AA and AB stacked BLG should take 2 years. 
The model fitting for benzene should take no more than four months as it is a reproduction of previous work.
The SLG model fitting should take another eight months.
The first two months will be dedicate to reproducing the on-site Hubbard model from that work.
Including long range Coulomb and electron-phonon interactions should take another four months.
The last two months will be geared towards finite size extrapolation to ensure the final model is applicable to bulk systems.
The last year of the project will be focused on the AA and AB BLG, with each project taking six months to complete.
For each stacking, I will need four months to build the model and two months for finite size extrapolation.
In summary, my work will contribute significantly to the understanding of low-energy effective interactions in single and bilayer graphene.

\vspace{-10pt}
\begin{multicols}{2}
\begin{spacing}{0.5}
{\footnotesize
\bibliography{biblio}
\bibliographystyle{prelim}
}
\end{spacing}
\end{multicols}
\end{document}